---
title: "Analysis of the Nile Dataset"
author: "Dipti Bhattad"
date: "February 11, 2019"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

The Nile dataset contains the time series data. Converting the time series data into a dataframe and adding an Year column to the dataframe
```{r}
nile_df <- as.data.frame(Nile)
nile_df$Year <- c(1871:1970)
```

Problem 1.1
Find the year when the river ???ow experienced a sudden change in the ???ow pattern for the ???rst time since 1871

The cpt.mean function in the changepoint library calculates the optimal positioning and number of changepoints for data
```{r}
#install.packages("changepoint")
library(changepoint)
cpmean=cpt.mean(nile_df$x)
cpmean
```
Thus, the year with the sudden change of flow is 1871+28= 1899


Problem 1.2
Plot the time series
```{r}
library('ggplot2')
ggplot(data = nile_df, aes(x = Year, y = x)) + 
  geom_line(color = "#FC4E07", size = 1) + ggtitle("Nile flow from 1871 to 1970") + ylab("Flow")
```


Problem 1.3
Draw a vertical line over the plot indicating the year of change as calculated from sub question 1
```{r}
ggplot(data = nile_df, aes(x = Year, y = x)) + 
  geom_line(color = "#FC4E07", size = 1) + 
  geom_vline(xintercept = (1871 + cpts(cpmean)), color = "blue", lty = 1, lwd = 1) +
  ggtitle("Nile flow from 1871 to 1970") +
  ylab("Flow")
```


Problem 1.4
Draw two trend lines, one from 1871 to the year when the river ???ow changed suddenly and the second from the point of change to 1970
```{r}
plot(cpmean, xlab = "Year", ylab = "Flow", main = "Nile flow from 1871 to 1970")
```

Alternate way of plotting Problem 1.4 using ggplot
```{r}
ggplot(data = nile_df, aes(x = Year, y = x)) + 
  geom_line(color = "#FC4E07", size = 1 ) + geom_smooth(data=subset(nile_df,nile_df$Year<=1899), method=lm,se=FALSE, color='black') +  geom_smooth(data=subset(nile_df,nile_df$Year>=1899),method=lm,se=FALSE,color='blue') + ggtitle("Nile flow from 1871 to 1970") + ylab("Flow")
```
