---
title: "Analysis and Data Visualization of Wine dataset"
author: "Dipti Bhattad"
date: "March 8, 2019"
output: html_document
---

### Reading the csv file
```{r}
wine_df <- read.csv("~/Desktop/winemag-data_first150k.csv")
```

### Loading the library and checking for null values
```{r}
library(tidyverse)
### check for Price null values
which(is.na(wine_df$price))
min(wine_df$price, na.rm=TRUE)
max(wine_df$price, na.rm=TRUE)

### check for Points null values
which(is.na(wine_df$points))
min(wine_df$points)
max(wine_df$points)
```
There are many rows with null values for price. However, not removing these null values as histogram automatically removes the null values.

There are no null values for points.


### Histogram of Price
```{r}
ggplot(data = wine_df) +
  geom_histogram(mapping = aes(x = price), fill = "light green", color = "dark green", alpha = .5, binwidth = 50) + xlab("Price") + ggtitle("Histogram of Price") + theme_classic()
```

### Histogram of Points
```{r}
ggplot(data = wine_df) +
  geom_histogram(mapping = aes(x = points), fill = "light pink", color = "tomato", alpha = .5, binwidth = 1) + xlab("Points") + ggtitle("Histogram of Points") + theme_classic()
```

### Create a new dataframe
```{r}
new_df <- summarise(group_by(wine_df,country,variety), avg_val_of_points = mean(points), avg_val_of_price = mean(price,na.rm=TRUE))
```

### Arrange the above dataframe in descending order of points. 
### Plot heatmap considering ???rst 10 values of the arranged dataframe. The x axis of heat map should have countries, y axis should have variety, and the heat map should be ???lled with values of average points 

Arranging the above dataframe in descending order of points and taking the first 10 records
```{r}
top_10_wines<- new_df %>% arrange(desc(new_df$avg_val_of_points)) %>% head(10)
```

Plotting the heatmap of the above dataframe
```{r}
ggplot(data = top_10_wines) + geom_tile(aes(x=country,y=variety, fill=avg_val_of_points)) + scale_fill_gradient2(low = "light green", mid = "orange", high = "red", midpoint = 94.5) + ggtitle("Heatmap of Varieties vs Countries based on avg points") + xlab("Countries") + ylab("Varieties") + theme_bw()
```


###  Compute the Pearson correlation coe???cient. Plot a scatter plot with average points on x axis and average price on y axis. Fit a linear model and compute the values of slope and intercept

Pearson correlation coeffecient
```{r}
cor(new_df$avg_val_of_points, new_df$avg_val_of_price, use = "complete.obs")
```
The Pearson correlation coeffecient is 0.49 which is NOT very high and hence does not indicate a strong relationship between average value of points and average value of price


Scatter plot of average value of price vs average value of points
```{r}
ggplot(data = new_df, mapping = aes(x = avg_val_of_points, y = avg_val_of_price)) + geom_point() + geom_smooth(method = "lm") + ggtitle("Scatter plot of avg price vs avg points") + theme_bw()
```
The Scatter plot also indicate that there is no strong relationship between average value of points and average value of price


Fitting a linear model
```{r}
linear_model <- lm(formula = avg_val_of_price ~ avg_val_of_points, data = new_df)
summary(linear_model)
```
Value of Intercept : -290.3801
Value of Slope: 3.6309


### Does a winery o???ering more variety of wines attract more points on average? Is there a trend between variety and average points?

```{r}
winery_df <- summarise(group_by(wine_df,winery), average_points = mean(points), total_varieties = n_distinct(variety))

ggplot(data = winery_df, mapping = aes(x = total_varieties, y = average_points)) + geom_point() + geom_smooth(method = "lm") + theme_bw() + ggtitle("Scatter plot of avg points vs total varieties for each winery")
```
No, the winery offering more varieties of wines do not attract more average points.
As seen from the above plot, there is no trend between variety and average points.
