---
title: "Network Visualization"
author: "Dipti Bhattad"
date: "March 8, 2019"
output: html_document
---

### Loading the libraries
```{r}
library(ggplot2)
library(igraph)
library(dplyr)
```

### Reading the edges file and creating a nodes dataframe
```{r}
links <- read.csv("~/Desktop/edgelist.csv", header=T, as.is=T)
links <- links[,2:3]
nodes <- sort(unique(links$to))
nodes <- as.data.frame(nodes)
net <- graph_from_data_frame(d=links, vertices=nodes, directed = TRUE)
net <- simplify(net, remove.multiple = TRUE, remove.loops = TRUE)
```


### Create an adjacency matrix
```{r}
adj_matrix <- as_adjacency_matrix(net, type = "both")
adj_matrix
```


### Create a network using the adjacency matrix
```{r}
net1<-graph_from_adjacency_matrix(adj_matrix,mode="undirected", diag = F)
#net1
plot(net1, edge.color="gray80",
     vertex.color="gold", vertex.frame.color="black", vertex.label.color="black", vertex.size = 25)
```


### Add a node (vertex) attribute and name it as type. Assign value A to nodes S01 to s08, value B to nodes S09 to s14, and value C to nodes S15 to s17. Plot the network showing all node labels and color node by type attribute

Creating a type attribute
```{r}
net1<-set.vertex.attribute(net, "type", value=c(rep(LETTERS[1], times = 8),rep(LETTERS[2], times = 6), rep(LETTERS[3], times = 3)))
V(net1)$type
```

Plotting the network
```{r}
V(net1)$color <- ifelse(V(net1)$type == "A", "tomato", ifelse(V(net1)$type=="B", "Green", "Gold"))

plot(net1, edge.color="gray80", vertex.frame.color="black", vertex.label.color="black", vertex.size = 25,vertex.color= V(net1)$color)
```


### Plot the network showing all nodes such that the node sizes are varied according their respective degrees

```{r}
colrs <- c("green", "pink", "yellow")

deg <- degree(net1, mode="all")
V(net1)$size <- deg*4

E(net1)$width <- E(net1)$weight/2
E(net1)$arrow.size <- .3
E(net1)$edge.color <- "gray"

graph_attr(net1, "layout") <- layout_with_lgl
plot(net1, edge.color="gray80", vertex.frame.color="black", vertex.label.color="black",vertex.color= colrs) 
```

